Draft_Application_Stochastic_Volatility

### 2021-07-19 Real Dataset SV.Rmd
commit

As regards the series that we will employ in section \textit{fill with the number of the section}, namely the series of realised volatilities which we will employ as a proxy for the true latent volatilities, the data were retrieved from the Oxford-Man Institute's Realized Library, which provides several different daily non-parametric measures of past volatility.\footnote{Such class of measures will be better presented in section \textit{fill with the number of the section}, when reviewing the use of realised volatility estimators as proxies for latent volatilities.} We consider a proper rescaling of the \textit{rv5} series, \textit{i.e.} model-free daily volatility estimates based on 5 min intraday return intervals. Also in this case, the chosen interval spanned from June 1st, 2017 to May 30th, 2021. \textit{Includere grafico e commento del grafico della realised volatility}

@@

We start by running the Bootstrap particle filter with N=50000, setting it as the benchmark. In table XX we report the RMSE and MAE measures one obtains by comparing the mean of the particles generated at each step by the Bootstrap PF with N=50000 and the mean of the particles generated by other particles filters with N equal to 10, 100, 1000, 5000 or 10000.\footnote{We consider 5 algorithms: a Bootstrap PF, a Guided PF (with the optimal proposal kernel), an Auxiliary PF, a "Basic PF" and a Sequential Importance Sampling alogrithm. By "Basic PF we refer to a Bootstrap PF that features resampling at each step, whereas resampling never happens in the SIS algorithm: one can then consider these 2 algorithms as extreme cases of an adaptive resampling algorithm, namely one that features resampling when ESS is non negative (\textit{i.e.} always) or when ESS<0 (\textit{i.e.} never). In this sense, we use the shorthand "particle filters" when referring to all the 5 algorithms.}

```{r include=FALSE}
Errorcomp1<-matrix(NA,ncol=6,nrow=10)
colnames(Errorcomp1)<-c("N","BPF", "GPFOPT", "APF", "BAPF","SIS")
rownames(Errorcomp1)<-c("RMSE","RMSE","RMSE","RMSE","RMSE","MAE","MAE","MAE","MAE","MAE")
Errorcomp1[,1]<-c(10,100,1000,5000,10000,10,100,1000,5000,10000)
Errorcomp1[,2]<-c(0.4007204,0.15268841,0.08482937,0.04275829,0.02931792,0.01369827,0.2688306,0.09413161,0.04579035,0.02021174)
Errorcomp1[,3]<-c(0.4123198,0.15152308,0.06776933,0.06018919,0.03052592,0.2757273,0.09076196,0.03282099,0.02622987,0.01399114)
Errorcomp1[,4]<-c(0.5002458,0.2493751,0.122905,0.06863069,0.04399511,0.3905036,0.1812517,0.08131335,0.03401982,0.02529053)
Errorcomp1[,5]<-c(0.4687276,0.2698023,0.12737504,0.08064348,0.0513611,0.3711974,0.1945553,0.07408736,0.03614264,0.02515361)
Errorcomp1[,6]<-c(0.9168639,1.1055065,0.6250516,0.6200359,0.5118927,0.7275327,0.9179852,0.4971839,0.4475265,0.406363)

@@

>>>>

As regards the series that we will employ in section \textit{fill with the number of the section}, namely the series of realised volatilities which we will employ as a proxy for the true latent volatilities, the data were retrieved from the Oxford-Man Institute's Realized Library, which provides several different daily non-parametric measures of past volatility.\footnote{Such class of measures will be better presented in section \textit{fill with the number of the section}, when reviewing the use of realised volatility estimators as proxies for latent volatilities.} We consider a proper rescaling of the \textit{rv5} series, \textit{i.e.} model-free daily volatility estimates based on 5 min intraday return intervals. Also in this case, the chosen interval spanned from June 1st, 2017 to May 30th, 2021. 
The plots below represent the realized volatility in the period of interest for all the three indices. As usual, they show a rather similar pattern across all three indices, identifying peaks corresponding to periods of higher uncertainty (in particular referring to the 2020 Covid-19 Crisis outbreak and its impact over time). Once more, there is great similarity between the series for the DJI and the one for the S&P500, while the series of the STOXX50E appears slightly different, in terms of peaks.

@@

We start by running the Bootstrap particle filter with N=50000, setting it as the benchmark. In table XX we report the RMSE and MAE measures one obtains by comparing the mean of the particles generated at each step by the Bootstrap PF with N=50000 and the mean of the particles generated by other particles filters with N equal to 10, 100, 1000, 5000 or 10000.\footnote{We consider 5 algorithms: a Bootstrap PF, a Guided PF (with the optimal proposal kernel), an Auxiliary PF, a "Basic PF" and a Sequential Importance Sampling alogrithm. By "Basic PF" we refer to a Bootstrap PF that features resampling at each step, whereas resampling never happens in the SIS algorithm: one can then consider these 2 algorithms as extreme cases of an adaptive resampling algorithm, namely one that features resampling when $ESS$ is non negative (\textit{i.e.} always) or when $ESS<0$ (\textit{i.e.} never). In this sense, we use the shorthand "particle filters" when referring to all the 5 algorithms.}

```{r include=FALSE}
Errorcomp1<-matrix(NA,ncol=6,nrow=10)
colnames(Errorcomp1)<-c("N","BPF", "GPFOPT", "APF", "BAPF","SIS")
rownames(Errorcomp1)<-c("RMSE","RMSE","RMSE","RMSE","RMSE","MAE","MAE","MAE","MAE","MAE")
Errorcomp1[,1]<-c(10,100,1000,5000,10000,10,100,1000,5000,10000)
Errorcomp1[,2]<-c(0.4007204,0.15268841,0.08482937,0.04275829,0.02931792,0.2688306,0.09413161,0.04579035,0.02021174,0.01369827)
Errorcomp1[,3]<-c(0.4123198,0.15152308,0.06776933,0.06018919,0.03052592,0.2757273,0.09076196,0.03282099,0.02622987,0.01399114)
Errorcomp1[,4]<-c(0.5002458,0.2493751,0.122905,0.06863069,0.04399511,0.3905036,0.1812517,0.08131335,0.03401982,0.02529053)
Errorcomp1[,5]<-c(0.4687276,0.2698023,0.12737504,0.08064348,0.0513611,0.3711974,0.1945553,0.07408736,0.03614264,0.02515361)
Errorcomp1[,6]<-c(0.9168639,1.1055065,0.6250516,0.6200359,0.5118927,0.7275327,0.9179852,0.4971839,0.4475265,0.406363)

@@

As we can see, almost every filter shows dramatic improvement already when passing from 10 particles to 100 particles, while the performance of the SIS algorithm actually gets worse. With N=10 or N=100, the Bootstrap PF and the Guided PF (with optimal proposal kernel) emerge quite clearly as the best performing filters; however, no conclusion should be drawn at this stage, as the large errors seem to suggest quite a wide margin from improvement even for the best performing filters.
